---
title: "Statistical Inference Course Project 1"
author: "Ilya Tishchenko"
date: "9/14/2020"

output:
  html_document: 
  
  toc: true
  
  toc_float: true
  
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```

## Part 1

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should:

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.


## Simulation

Generate thousand simulations by 40.

```{r simulation}

lambda <- 0.2
n <- 40
ns <- 1000
exp=NULL
for (i in 1 : 10000) exp = c(exp, mean(rexp(40,0.2)))
hist(exp)
```

```{r}
1/0.2 #theoretical mean

mean(exp)#sample mean

```

As could be seen theoretical and sample means are very close.  
  
## Variance comparison  
Theoretical standard deviation: $s=\frac{1/\lambda}{\sqrt{n}}$
```{r}
1/0.2/sqrt(40) #theoretical sd
sd(exp) #sample sd
```

Variance: $\sigma=s^2$
```{r}
(1/0.2/sqrt(40))^2 #theoretical variance
var(exp) # sample variance
```

## Show that distribution is approximately normal

In a same manner we generate normal distribution with appropriate mean and sd.
```{r normal dist}
# norm.sim=NULL
# for (i in 1 : 10000) norm.sim <- c(norm.sim,rnorm(40,5,5))
norm.sim <- rnorm(10000,5,5)

```
I plot density plot for both distributions exponential and normal, as could be seen visually the distribution of exponential simulation follows that of normal.
```{r}
#merge data into dataframe
exp.df <- data.frame(exp) %>%
  mutate(label="exp")%>%
  rename(value="exp")
norm.df <- data.frame(norm.sim)%>%
  mutate(label="norm")%>%
  rename (value="norm.sim")
sim.df <- rbind(exp.df,norm.df)

#calc CI for both dataset
cinorm <- quantile(norm.sim,c(0.025,.975))
ciexp <- quantile (exp,c(0.025,0.975))

cinorm.df <- data.frame(cinorm)%>%
  rename (value="cinorm")%>%
  mutate(label="norm")
ciexp.df <- data.frame(ciexp) %>%
  rename(value="ciexp")%>%
  mutate(label="exp")

ci.df <- rbind(cinorm.df, ciexp.df)

#plot histogramm, mean and CI
ggplot(sim.df)+
  geom_density(aes(x=value,fill=label), alpha=.5)+
  labs(title="Density plot for two simulated vectors",subtitle = " with mean shown as dashed and CI±95% as solid lines", x="values")+
  scale_fill_discrete(name="", labels=c("Exponential","Normal"))+
  geom_vline(aes(xintercept=mean(value),color=label),size=1,linetype=2, show.legend = FALSE)+
  geom_vline(data=ci.df,aes(xintercept=value,color=label), show.legend = FALSE)+
  theme_minimal()
  
```

